### 2021-09-13 note-temp/note-cover.tex

% Note Cover

\begin{titlepage}
\begin{tikzpicture}[remember picture, overlay]
\fill[cyan!20](current page.north west) rectangle (current page.south east);
\fill[hong!90](current page.north west) rectangle ([xshift=0.8cm,yshift=-7cm]current page.north west);
\fill[hong!50]([yshift=-7cm]current page.north west) rectangle ([xshift=0.8cm,yshift=-11cm]current page.north west);
\draw[line width=2pt,huang]([yshift=-7cm]current page.north west)--node[above=20pt]{\color{black}\HUGE \newtitle}([yshift=-7cm]current page.north east);
\node[opacity=0.25,rotate=30] at ([xshift=4.8cm,yshift=-3cm]current page.north west){\LARGE $e^{\mathrm{i}\pi}+1=0$};
\node[circle,fill,inner sep=5pt]at([xshift=10.7cm,yshift=-8.1cm]current page.north west)
{};
\node[dart,fill=white,inner sep=2pt]at([xshift=10.7cm,yshift=-8.1cm]current page.north west){};
\node[right] at ([xshift=11.2cm,yshift=-8.1cm]current page.north west){\Large\fangsong 机构名称 \quad 作者 \quad 编写};
\node[opacity=0.45,right] at ([xshift=3.5cm,yshift=-22cm]current page.north west)
{\includegraphics[scale=1.6]{titlefig01.png}};
\node[opacity=0.45,right] at ([xshift=10.5cm,yshift=-13cm]current page.north west)
{\includegraphics[scale=2.4]{titlefig02.png}};
\end{tikzpicture}
\end{titlepage}