### 2022-10-29 note-temp/matlab-style.tex

%%%%%%%%%%%%%%%%%%%%%%%%% Code Style %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Matlab highlight color settings
%\definecolor{mBasic}{RGB}{248,248,242}       % default
\definecolor{mKeyword}{RGB}{0,0,255}          % bule
\definecolor{mString}{RGB}{160,32,240}        % purple
\definecolor{mComment}{RGB}{34,139,34}        % green
\definecolor{mBackground}{RGB}{245,245,245}   % lightgrey
\definecolor{mNumber}{RGB}{128,128,128}       % gray

\lstset{language=matlab,               % choose the language of the code
  xleftmargin=20pt,
  xrightmargin=20pt,
  frame=tlbr,framesep=6pt,framerule=0.5pt, % sets the frame style
  %frame=shadowbox,rulesepcolor=\color{red!20!green!20!blue!20},
  basicstyle=\footnotesize\ttfamily,
  keywordstyle={\color{mKeyword}},     % sets color for keywords
  stringstyle={\color{mString}},       % sets color for strings
  commentstyle={\color{mComment}},     % sets color for comments
  backgroundcolor=\color{gray!5},      %\color{pBackground}, % choose the background color
  rulecolor=\color{gray!50},
  %title=\lstname,                      % \lstname show the filename of files
  keywords={break,case,catch,classdef,continue,else,elseif,end,for,
  function,global,if,otherwise,parfor,persistent,return,spmd,switch,try,while},
  showspaces=false,                    % show spaces adding particular underscores
  showstringspaces=false,              % underline spaces within strings
  showtabs=false,                      % show tabs within strings adding particular underscores
  tabsize=4,                           % sets default tabsize to 2 spaces
  captionpos=t,                        % sets the caption-position to bottom
  breaklines=true,                     % sets automatic line breaking
  numberstyle=\tiny\color{mNumber},
  numbers=left,                        % where to put the line-numbers
  stepnumber=1,                        % the step between two line-numbers.
}

%\lstinputpath{./M-files/}
\newcommand*\lstinputpath[1]{\lstset{inputpath=#1}}
\lstinputpath{{./M-files/}}

\lstset{basicstyle=\footnotesize\fontspec{Consolas}}
%\lstset{basicstyle=\footnotesize\fontspec{Courier New}}

%\renewcommand{\lstlistingname}{\texttt{Listing}}%
%
%\newcommand{\mylisting}[2][]{%
%    \lstinputlisting[style=MyStyle,caption=\texttt{\detokenize{#2}},#1]{#2}%
%}

%\usepackage[numbered,autolinebreaks,useliterate]{mcode} % framed,
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%