### 2023-02-13 amsart-en/amsart-temp.tex
update

% paper template

\documentclass[10pt,reqno,final]{amsart}
%\documentclass[10pt,reqno,draft]{amsart}

% package for mathematics
\usepackage{amsmath,amssymb,amsthm}
% package for RSFS fonts in maths
\usepackage{mathrsfs}
% package for special symbols
\usepackage{pifont}
% package for multilingual support
\usepackage[english]{babel}
% package for figures
\usepackage{graphicx}
\usepackage{subfigure}
% package for hyperlink
\usepackage{url,hyperref}
% package for layout of list
\usepackage{enumitem}
% package for showing keys in draft mode
\usepackage[notcite,notref]{showkeys}
% package for color
\usepackage{color}
% package for table
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{makecell}
% package for multiple columns and rows
\usepackage{multirow,multicol}
% package for caption
\usepackage{caption}
% package for algorithm
\usepackage{algorithm}
\usepackage{algpseudocode}
% package for using box and verbatim
\usepackage{fancybox}

% allow page breaks between multiline formulas
\allowdisplaybreaks

% layout setting
\setlength{\textwidth}{15cm}
\setlength{\textheight}{21.6cm}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}

% command for line spacing
%\renewcommand{\baselinestretch}{1.1}

% command for mark footnotes
%\usepackage[symbol]{footmisc}
%\renewcommand{\thefootnote}{\fnsymbol{footnote}}

% command for equations, theorems and lemmas etc.
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}
\newtheorem{xca}{Exercise}[section]

\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

% caption setting
%\captionsetup{font={small,singlespacing},labelformat={default},labelsep=period}
\captionsetup[figure]{position=bottom,skip={5pt},name={Figure}}
\captionsetup[table]{position=top,skip={2pt},name={Table}}
%\setlength{\textfloatsep}{12pt plus 2pt minus 2pt}
%\setlength{\floatsep}{10pt plus 2pt minus 2pt}
%\setlength{\intextsep}{10pt plus 2pt minus 2pt}
%\setlength{\abovecaptionskip}{2pt plus 1pt minus 1pt}
%\setlength{\belowcaptionskip}{3pt plus 1pt minus 2pt}

% number of equation, figure and table
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}

% new command
\newcommand{\abs}[1]{\lvert#1\rvert}
\newcommand{\dx}[1][x]{\,{\rm d}#1}
\newcommand{\ddx}[1][x]{\dfrac{\rm d}{{\rm d}#1}}
\newcommand{\refe}[2]{(\ref{#1})--(\ref{#2})}
\newcommand{\red}[1]{\textcolor{red}{#1}}

% Information for title and author
\title[Short Title]{This is Full Title}

\author[Author A \; \& \; Author B]{Author A${\,}^{1}$ \quad and \quad Author B${\,}^{2,*}$}

\thanks{${}^{*}$ Corresponding author (E-mail: \texttt{xyz@math.univ.edu}). \\
\indent ${}^{1}$ Department of Computer Science, \LaTeX\ University. \\
\indent ${}^{2}$ Department of Mechanical Engineering, \LaTeX\ University \\
\indent This work is supported in part by NSFC grants No. 12345678 and ...
}

\date{\today}

\subjclass[2000]{65L60, 65L05, 65L70}

\keywords{Keyword 1, keyword 2, error analysis}

\begin{document}

\begin{abstract}
  This is an example \LaTeX\ article. This can be used as a
  template for new articles.  Abstracts must be able to stand alone
  and so cannot contain citations to the paper's references,
  equations, etc.  An abstract must consist of a single paragraph and
  be concise. Because of online formatting, abstracts must appear as
  plain as possible. Any equations should be inline.
\end{abstract}

\maketitle

\section{Introduction}
The introduction introduces the context and summarizes the
manuscript. It is importantly to clearly state the contributions of
this piece of work.

In this paper we present a new method for solving the model equation
\begin{equation}\label{eq:mulequ}
\left\{\begin{aligned}
  & \partial_{t} u-\varepsilon^{2} \Delta u+u^{3}-u=0, \quad \text{ in } ~\Omega\times\mathcal{T}, \\
  &\, u(x,y,t) = g(t), \quad \text{ on } ~ \partial \Omega, \\
  &\, u(x,y,0)=\varphi(x, y), \quad \text{ on } ~\Omega.
\end{aligned}\right.
\end{equation}
where $\varepsilon$ is a small parameter.

This is an example of quoting an equation \eqref{eq:mulequ}.

The merits of our method are as follows:
\begin{itemize}
  \item item one
  \item item two
  \item item three
\end{itemize}
\begin{enumerate}
  \item item one
  \item item two
  \item item three
\end{enumerate}

The outline is not required, but we show an example here.

The paper is organized as follows. Our main results are in
\ref{sec:main}, our new algorithm is in \ref{sec:alg}, experimental
results are in \ref{sec:experiments}, discussion is in \ref{sec:discussion} and the conclusions follow in
\ref{sec:conclusions}.

\section{Main results}
\label{sec:main}

%\subsection{Subsection name}

We interleave text filler with some example theorems and theorem-like
items.

\begin{definition}
This is a definition environment.
\end{definition}

\begin{lemma}
This is a lemma environment.
\end{lemma}

\begin{theorem}
This is a theorem environment.
\end{theorem}
\begin{proof}
  This is a proof environment.
\end{proof}

\begin{lemma}
This is a lemma environment
\begin{enumerate}[label=\rm (\roman*)]
  \item item A
  \item item B
  \begin{equation}\label{eq:limite}
    \lim_{n\to\infty}\left(1+\frac{1}{n}\right)^n=e.
  \end{equation}
\end{enumerate}
\end{lemma}

This is a citation example \cite{WoZhMeSh2005}.

Here we state our main result as \ref{thm:bigthm}.

\begin{theorem}[$LDL^T$ Factorization \cite{GoVa2013}]\label{thm:bigthm}
  If $A \in \mathbb{R}^{n \times n}$ is symmetric and the principal
  submatrix $A(1:k,1:k)$ is nonsingular for $k=1:n-1$, then there
  exists a unit lower triangular matrix $L$ and a diagonal matrix
  \begin{equation*}
    D = \operatorname{diag}(d_1,\dots,d_n),  %\diag(d_1,\dots,d_n)
  \end{equation*}
  such that $A=LDL^T$. The factorization is unique.
\end{theorem}

LaTeX is a high-quality typesetting system; it includes features designed
for the production of technical and scientific documentation.
LaTeX is the de facto standard for the communication and publication of scientific documents.
LaTeX is based on the idea that it is better to leave document design to
document designers, and to let authors get on with writing documents.

\begin{theorem}[Mean Value Theorem]\label{thm:mvt}
  Suppose $f$ is a function that is continuous on the closed interval
  $[a,b]$.  and differentiable on the open interval $(a,b)$.
  Then there exists a number $c$ such that $a < c < b$ and
  \begin{equation*}
    f'(c) = \frac{f(b)-f(a)}{b-a}.
  \end{equation*}
  In other words,
  \begin{equation*}
    f(b)-f(a) = f'(c)(b-a).
  \end{equation*}
\end{theorem}

\begin{remark}
Observe that \ref{thm:bigthm}, \ref{thm:mvt} correctly mix references
to multiple labels.
\end{remark}

% command cref package cleveref
%Observe that \cref{thm:bigthm,thm:mvt,cor:a} correctly mix references
%to multiple labels.

\begin{corollary}\label{cor:a}
  Let $f(x)$ be continuous and differentiable everywhere. If $f(x)$
  has at least two roots, then $f'(x)$ must have at least one root.
\end{corollary}
\begin{proof}
  Let $a$ and $b$ be two distinct roots of $f$.
  By \ref{thm:mvt}, there exists a number $c$ such that
  \begin{equation*}
    f'(c) = \frac{f(b)-f(a)}{b-a} = \frac{0-0}{b-a} = 0.
  \end{equation*}
\end{proof}

Note that it may require two \LaTeX\ compilations for the proof marks
to show.

Display matrices can be rendered using environments from \texttt{amsmath}:
\begin{equation}\label{eq:matrices}
S=\begin{bmatrix}1&0\\0&0\end{bmatrix}
\quad\text{and}\quad
C=\begin{pmatrix}1&1&0\\1&1&0\\0&0&0\end{pmatrix}.
\end{equation}
Equation \ref{eq:matrices} shows some example matrices.

We calculate the Fr\'{e}chet derivative of $F$ as follows:
\begin{subequations}
\begin{align}
  F'(U,V)(H,K)
  &= \langle R(U,V),H\Sigma V^{T} + U\Sigma K^{T} -
  P(H\Sigma V^{T} + U\Sigma K^{T})\rangle \label{eq:aa} \\
  &= \langle R(U,V),H\Sigma V^{T} + U\Sigma K^{T}\rangle
  \nonumber \\
  &= \langle R(U,V)V\Sigma^{T},H\rangle +
  \langle \Sigma^{T}U^{T}R(U,V),K^{T}\rangle. \label{eq:bb}
\end{align}
\end{subequations}
\ref{eq:aa} is the first line, and \ref{eq:bb} is the last line.

\clearpage
\section{Algorithm}
\label{sec:alg}

Our analysis leads to the algorithm in \ref{alg:buildtree}.

\begin{algorithm}
\caption{Build tree}
\label{alg:buildtree}
\begin{algorithmic}
  \State {Define $P:=T:=\{ \{1\},\ldots,\{d\}$\}}
  \While{$\#P > 1$}
  \State {Choose $C^\prime\in\mathcal{C}_p(P)$ with $C^\prime := \operatorname{argmin}_{C\in\mathcal{C}_p(P)} \varrho(C)$}
  \State {Find an optimal partition tree $T_{C^\prime}$ }
  \State {Update $P := (P{\setminus} C^\prime) \cup \{ \bigcup_{t\in C^\prime} t \}$}
  \State {Update $T := T \cup \{ \bigcup_{t\in\tau} t : \tau\in T_{C^\prime}{\setminus} \mathcal{L}(T_{C^\prime})\}$}
  \EndWhile
  \Return $T$
\end{algorithmic}
\end{algorithm}

Reduce the width of the algorithm environment
\begin{center}
\vspace{-2ex}
\begin{minipage}{0.9\linewidth}
\begin{algorithm}[H]
\caption{Euclid’s algorithm}
\label{euclid}
\begin{algorithmic} % The number tells where the line numbering should start
    \Procedure{Euclid}{$a,b$} \Comment{The g.c.d. of a and b}
        \State $r\gets a \bmod b$
        \While{$r\not=0$} \Comment{We have the answer if r is 0}
            \State $a \gets b$
            \State $b \gets r$
            \State $r \gets a \bmod b$
        \EndWhile\label{euclidendwhile}
            \State \textbf{return} $b$\Comment{The gcd is b}
    \EndProcedure
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

\section{Experimental results}
\label{sec:experiments}

LaTeX is a high-quality typesetting system; it includes features designed
for the production of technical and scientific documentation.
LaTeX is the de facto standard for the communication and publication of scientific documents.
LaTeX is based on the idea that it is better to leave document design to
document designers, and to let authors get on with writing documents.

\begin{example}
  This is example environment.
\end{example}

Figure \ref{fig:a} shows some example results.
\begin{figure}[htp!]
  \centering
  \includegraphics[width=0.48\linewidth]{fig1}
  \caption{Example figure using external image files.}
  \label{fig:a}
\end{figure}

\clearpage
The two figures are placed side by side, sharing the same title.
\begin{figure}[htp!]
  \centering
  \includegraphics[width=0.45\linewidth]{fig1}
  \hfill
  \includegraphics[width=0.45\linewidth]{fig2}
  \caption{Left: Caption 1, Right: Caption 2.}
  \label{fig:b}
\end{figure}

Use \texttt{minipage} package to set up images side-by-side, each with its own title.
\begin{figure}[htp!]
\begin{minipage}[t]{0.48\linewidth}
  \centering
  \includegraphics[width=0.96\linewidth]{fig1}
  \caption{Caption A}
  \label{fig:image1}
\end{minipage}
\hfill
\begin{minipage}[t]{0.48\linewidth}
\centering
   \includegraphics[width=0.95\linewidth]{fig2}
   \caption{Caption B}
   \label{fig:image2}
\end{minipage}
\end{figure}

Additional results are available in the supplement in Table \ref{tab:foo}.

\begin{table}[!htp]
\centering
\renewcommand\arraystretch{1.2} % line spacing
\caption{Numerical error}
\label{tab:foo}
\begin{tabular}{c|c|cc|cc|cc}
\hline
degree &  step-size~$h$  & $L^2$-errors  &  order  & $H^1$-errors & order & $L^\infty$-errors  &  order \\
\hline
   &  1/128    & 9.18E-06    &2.02    & 7.70E-03  &1.01       & 6.46E-07    &2.02   \\
1  &  1/256    & 2.29E-06    &2.01    & 1.92E-03  &1.00       & 1.61E-07    &2.01   \\
   &  1/512    & 5.70E-07    &2.00    & 9.56E-04  &1.00       & 4.01E-08    &2.00   \\
\hline  % \cline{1-8}
   &  1/128    & 1.39E-08    &3.01    & 1.15E-05  &2.01       & 3.48E-12   &4.02    \\
2  &  1/256    & 1.73E-09    &3.01    & 2.88E-06  &2.01       & 3.27E-13   &3.94    \\
   &  1/512    & 2.17E-10    &3.00    & 7.24E-06  &2.00       & 6.66E-13   &1.55    \\
\hline  % \cline{1-8}
   &  1/32     & 2.28E-09    &4.05    & 6.92E-07  &3.04       & 1.45E-15   &8.21    \\
3  &  1/64     & 1.42E-10    &4.03    & 8.65E-08  &3.02       & 2.06E-14   &3.85    \\
   &  1/128    & 8.91E-12    &4.01    & 1.08E-08  &3.01       & 3.86E-14   &0.91    \\
\hline
\end{tabular}
\end{table}

\clearpage
Use the \texttt{tabularx} environment to generate tables.
\begin{table}[htp!]
\centering
\newcolumntype{L}{X}
\newcolumntype{C}{>{\centering \arraybackslash}X}
\newcolumntype{R}{>{\raggedleft \arraybackslash}X}
\newcolumntype{P}[1]{>{\centering \arraybackslash}p{#1}}
\renewcommand\arraystretch{1.2}
\caption{Table description}
\label{tab:error}
\begin{tabularx}{0.8\textwidth}{|P{0.8cm}|C|C|C|C|C|}
\Xhline{2\arrayrulewidth}
N  & A       & B    & C       & D      & E     \\
\Xhline{2\arrayrulewidth}
2  & 9.20E-05 & 9.90E-05 & 1.00E-06 & 8.00E-06 & 1.50E-05 \\
4  & 9.80E-05 & 8.00E-05 & 7.00E-06 & 1.40E-05 & 1.60E-05 \\
6  & 4.00E-06 & 8.10E-05 & 8.80E-05 & 2.00E-05 & 2.20E-05 \\
8  & 8.50E-05 & 8.70E-05 & 1.90E-05 & 2.10E-05 & 3.00E-06 \\
10 & 8.60E-05 & 9.30E-05 & 2.50E-05 & 2.00E-06 & 9.00E-06 \\
12 & 1.70E-05 & 2.40E-05 & 7.60E-05 & 8.30E-05 & 9.00E-05 \\
\Xhline{2\arrayrulewidth}
\end{tabularx}
\end{table}

\section{Discussion of \texorpdfstring{{\boldmath$Z=X \cup Y$}}{Z = X union Y}}
\label{sec:discussion}

Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here. Some discussions here.

\section{Conclusions}
\label{sec:conclusions}

Some conclusions here.

\appendix
\section{An example appendix}

The contents of the appendix are here.

\begin{lemma}
Test Lemma.
\end{lemma}

This is a equation in appendix.
\begin{equation}\label{eq:A1}
  a^2+b^2=c^2.
\end{equation}

\section*{Acknowledgments}
We would like to acknowledge the assistance of volunteers in putting
together this example manuscript and supplement.

% References
\bibliographystyle{plain}
%\bibliographystyle{abbrv}

\bibliography{reference}

\end{document}