### 2023-05-09 cover-letter/Cover-Letter.tex
update

% LaTeX cover letter template

\documentclass[11pt]{article}

% all kinds of packages needed
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{bm,mathrsfs}
\usepackage{graphicx}
\usepackage{color,xcolor}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{mathtools}
%\usepackage{charter} % Charter font
\usepackage{geometry}
\geometry{left=1.25in,right=1.25in,top=1in,bottom=1in}

% Default font family: sans serif
%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}
\pagenumbering{gobble} % No page numbering

% allow page breaks between multiline formulas
\allowdisplaybreaks

\setlength{\parindent}{0pt}  % Paragraph indentation
\setlength{\parskip}{1em}    % Vertical space between paragraphs

%-----------------------------------------------------

\begin{document}

%-----------------------------------------------------
%	First page header
%-----------------------------------------------------

%\includegraphics[width=0.5\textwidth]{UnivLogo.png} % Logo

%\vspace{-1em}

%\rule{\linewidth}{0.5pt} % Horizontal rule

\vskip2.5em

%-----------------------------------------------------
%	Your information
%-----------------------------------------------------

\hfill
\begin{tabular}{l @{}}
	\textbf{XX research group} \\
    \textbf{Department of Mathematics} \\
    University of XXXX \\
    XX Building, XX Road \\
	City 200123, P. R. China. \\[10pt] % Address
    \textbf{Dr. Who} \\
	Email: xxx@univ.edu.cn \bigskip \\
    \today \\ % Date
\end{tabular}

\bigskip

\bigskip

%Dear Mrs.\ XXX,
Dear Editor,

\bigskip

%-----------------------------------------------------
%	Letter body
%-----------------------------------------------------

Please find attached the manuscript ``PAPER TITLE" by Author A and Author B for consideration as a Research Article in XX.

Our work builds on the considerable interest of the scientific community in XXX [cite]. This  is more impressing especially since the XX were recently experimentally demonstrated [cite].

Hitherto, XX have increasingly been investigated [cite]. However, the ``missing component" has  been XXX. XXX will enable future XXX.

In this paper, we XXX.

We are convinced that this work will have implications for XXX. As such, the research should  appeal to a broad audience interested in XXX. We thus believe that the manuscript is  appropriate for the readership of XX.

We remain at your disposal for any further questions.

\bigskip %

Sincerely yours,

\includegraphics[height=1.5cm]{signature}

%\vspace{25pt} %

Andy

\end{document}