### 2023-05-09 response-letter/Review-Report.tex

% Review report

\documentclass{article}
\usepackage{amsthm,amsmath,amssymb}
\usepackage{bm,mathrsfs}
\usepackage{graphicx}
\usepackage{color,xcolor}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{mathtools}
\usepackage{geometry}
\geometry{left=1.25in,right=1.25in,top=1.25in,bottom=1.25in}

\renewcommand{\baselinestretch}{1.5}
\pagenumbering{gobble}
%\pagestyle{plain}

% information of paper
\title{Review Report on \\
``This is the title of the journal article"
}
\author{by Reviewer}
\date{}

\begin{document}

\maketitle

This article is devoted to developing a new method for the model equation.
This manuscript aims to give upper and lower bounds for the posterior error estimate.
The goal of this manuscript is to provide comprehensive information on the topic.
The results obtained in this manuscript are very meaningful, providing valuable insights into the development of the XXX method for the model equation.
The main contributions of this paper are: (1) The first innovation of the paper; (2) The second innovation of the paper.

The reviewer has the following comments.

\begin{enumerate}[label={\arabic*.}]
\item Page 2, Line 12: ``Expression A" should be ``Expression B";
\item Page 3, Line 5: The notation ``$u(t)$" should be ``$\tilde{u}(t)$";
\item Page 5, Line 8: The value ranges for the parameter $C$ in equation (2.1) is not clear, please give a little more detail about it.
\item Page 8, Line 4: It should be added a comma.
\item Page 10, Line $-2$:  The function $f(x)=b$ contradicts with the definition of $f(x)=a$.
\end{enumerate}

\end{document}