### 2024-04-12 poster/poster1-en.tex

%======================================================%
%   Poster LaTeX Template
%   compile using PDFLaTeX
%======================================================%

\documentclass{beamer}
\usepackage{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
%\usepackage[UTF8,indent]{ctex}
%\usepackage{indentfirst}
\renewcommand{\indent}{\hspace*{2em}}

\usepackage{microtype}

\bibliographystyle{plain}
\usepackage[orientation=portrait,size=a4,scale=2.3]{beamerposter} %1.68

\setbeamertemplate{navigation symbols}{}
\setbeamersize{text margin left=1.5cm, text margin right=1.5cm}

\renewcommand{\baselinestretch}{1.15}

%\usetheme{LLT-poster}
%\usecolortheme{ComingClean}
%\footimage{\includegraphics[width=.35\textwidth]{authors}}

% Default font family: sans serif
%\usepackage[no-math]{fontspec}
%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}

\makeatletter
\setbeamertemplate{frametitle}{%
  \nointerlineskip%
  \vskip0.4cm%
  \begin{beamercolorbox}[wd=\paperwidth,leftskip=.3cm,rightskip=.3cm plus1fil,vmode]{frametitle}%
  \hspace*{4pt} \includegraphics[height=1.5cm]{\logopic} %Logo.png
  \begin{center}
    \usebeamerfont{frametitle}
    \insertframetitle
  \end{center}
  \end{beamercolorbox}%
  %\vskip-6pt
  \begin{beamercolorbox}[wd=\paperwidth,leftskip=.3cm,rightskip=.3cm plus1fil,center]{title in head/foot}%
    \usebeamerfont{title in head/foot}
    \parbox[t]{15cm}{\centering \inserttitle}
  \end{beamercolorbox}%
  \vskip6pt
  \begin{beamercolorbox}[wd=\paperwidth,leftskip=.3cm,rightskip=.3cm plus1fil,center]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertauthor
  \end{beamercolorbox}%
  \vskip-8pt
  \begin{beamercolorbox}[wd=\paperwidth,leftskip=.3cm,rightskip=.3cm plus1fil,center]{author in head/foot}%
    \usebeamerfont{institute in head/foot}\insertinstitute
  \end{beamercolorbox}%
  \begin{beamercolorbox}[colsep=1.5pt,wd=14cm,leftskip=.3cm,rightskip=.3cm]{lower separation line head}
  \end{beamercolorbox}
}
\makeatother

% needed to define \columnheight
\newlength\columnheight \columnheight=5cm%
\def\postercolumn{
    \leavevmode
    \vtop to \columnheight\bgroup
    \hsize.5\textwidth
    \noindent
    \ignorespaces
}
\def\endpostercolumn{
    \egroup
}

% set fonts
\setbeamerfont{frametitle}{family=\rmfamily,size=\LARGE,series=\bfseries}
\setbeamerfont{title in head/foot}{family=\rmfamily,size=\Large}
\setbeamerfont{author in head/foot}{family=\rmfamily,size=\large,series=\bfseries}
\setbeamerfont{institute in head/foot}{family=\rmfamily,size=\large,series=\bfseries}

\setbeamercolor{frametitle}{fg=black}
\setbeamercolor{author in head/foot}{fg=black}

% set block font
\setbeamerfont{block title}{family=\rmfamily,series=\bfseries}
\setbeamerfont{block body}{family=\rmfamily}
\setbeamercolor{block title}{fg=black}
%\setbeamercolor{block body}{size=black}

\setbeamerfont{normal text}{family=\rmfamily}
\AtBeginDocument{\usebeamerfont{normal text}}

\def\logopic{UnivLogo.png}  % logo
\title{This is a LaTeX template of poster}
\author{Professor ~ John Smith}
\institute{Department of Mathematics}

\begin{document}

% only change the spacing of body text
%\setlength{\baselineskip}{16pt}

\begin{frame}
\frametitle{Academic Poster}

\vskip-5em

\begin{columns}[t] %c

\column{0.75\textwidth}

\begin{block}{About the Speaker}
\indent This is a brief introduction to the author. This is a brief introduction to the author.
This is a brief introduction to the author. This is a brief introduction to the author.
This is a brief introduction to the author. This is a brief introduction to the author.
This is a brief introduction to the author.

\end{block}

\column{0.1\textwidth}
\vspace{4ex}
\begin{center}
\hspace{-6ex}\includegraphics[width=1.5\textwidth]{author.jpg}
\end{center}

\end{columns}

%\begin{postercolumn}
%   hello world hello world hello world hello world
%   hello world hello world hello world hello world
%\end{postercolumn}

\vskip1em

\begin{block}{Abstract}
\indent This is abstract. This is abstract. This is abstract. This is abstract. This is abstract. This is abstract. This is abstract. This is abstract. This is abstract. This is abstract. This is abstract. This is abstract.
\end{block}

\vskip1em

{\bfseries\large Date:} October 28, 2022 (Friday)
\medskip

{\bfseries\large Time:} 9:00--10:00
\medskip

{\bfseries\large Venue:} Room 001, Building 3
\medskip

{\bfseries\large Tencent Meeting:} 111-222-333
\medskip

\vspace{1em}

\begin{center}
\LARGE\bfseries Welcome!
\end{center}

\vspace{1em}

\begin{flushright}
Department of Mathematics \\
XX University  \\
October 26, 2022
\end{flushright}

\end{frame}

\end{document}